\chapter*{Abstract}
\addcontentsline{toc}{chapter}{Abstract}

Application modernization is a major challenge for organizations operating systems that rely on outdated versions of programming languages, frameworks, and dependencies. Despite the availability of specialized tools, a substantial part of software migration remains manual: understanding the source project, identifying incompatibilities, transforming the codebase, executing builds and tests, and repeating diagnostic and correction cycles. This situation increases delivery time, cost, and reliance on specialized technical expertise.

This final-year engineering project, conducted at \textbf{CGI Technologies and Solutions Morocco}, focuses on the design of an \textbf{intelligent, extensible, and governed Migration Factory}. Its objective is to industrialize the migration process by combining a deterministic foundation with specialized agents and large language models. The deterministic layer encapsulates the tools, commands, rules, and validations required by each technology ecosystem. It retains authority over workflow state, phase transitions, execution within isolated workspaces, and the collection of logs and technical evidence.

Agents and large language models complement this foundation by accelerating activities that require deeper analysis and adaptation. They support project analysis, migration-plan enrichment, result interpretation, and execution monitoring, while having no direct authority to apply changes. When an installation, build, or test failure occurs, the system constructs a repair context from the detected errors and the current pipeline state. A model then proposes a corrective change, which is independently reviewed and submitted to the developer for approval before application and revalidation.

The implementation is organized into two separate software repositories: one dedicated to \textbf{Java/Spring Boot} migration and the other to \textbf{Angular} migration. This separation preserves the technical specificities of each ecosystem, avoids unnecessarily complex orchestration, and facilitates the evolution, testing, and long-term maintenance of each solution. Both repositories nevertheless share the same principles of governance, traceability, isolation, and controlled collaboration between deterministic automation, large language models, and human expertise.

The main contribution is therefore a controlled approach that aims to reduce manual effort, accelerate migration and repair cycles, improve operational reproducibility, and make every decision auditable, without pursuing full autonomy at the expense of safety and human oversight.

\vspace{0.5cm}
\noindent\textbf{Keywords:} Software migration, application modernization, Migration Factory, deterministic automation, generative artificial intelligence, software agents, large language models, Java, Spring Boot, Angular, orchestration, traceability, human-in-the-loop, software repair.